---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)

options(tibble.print_min = 5, tibble.print_max = 5)

set.seed(27)
```

# safepredict
[![Travis build status](https://travis-ci.org/alexpghayes/safepredict.svg?branch=master)](https://travis-ci.org/alexpghayes/safepredict)
[![Coverage status](https://codecov.io/gh/alexpghayes/safepredict/branch/master/graph/badge.svg)](https://codecov.io/github/alexpghayes/safepredict?branch=master)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

## Development notes

Working towards a MVP at the moment. NO GUARANTEES ABOUT PREDICTIONS yet other than that you get something back, and it's probably in a tibble. When implementing new methods, be sure to look into the helpers:

- `as_pred_tibble()` (S3 generic) and accompanying `binomial/multinomial_helper()`

- `multi_predict()` code has been copy-pasted over from `parsnip` but is a low priority at the moment. the `multi_predict()` interface is mostly settled
- many `type`s are missing and need to be filled in
- tests is quite limited

The goal of safepredict is to provide consistent predictions via the `safe_predict()` generic. `safe_predict()`:

- always returns a tibble.
- never drops rows with missing data.
- follows a consistent naming convention.

## Installation

`safepredict` is currently in the beginning stages of development and is available only on Github. You can install it with:

```r
# install.packages("devtools")
devtools::install_github("alexpghayes/safepredict")
```

## Examples

Suppose you fit a logistic regression using `glm`:

```{r}
library(tibble)

data <- tibble(
  y = as.factor(rep(c("A", "B"), each = 50)),
  x = c(rnorm(50, 1), rnorm(50, 3))
)

fit <- glm(y ~ x, data, family = binomial)
```

You can predict class probabilities:

```{r}
library(safepredict)

test <- tibble(x = rnorm(10, 2))

safe_predict(fit, new_data = test, type = "prob")
```

or can jump straight to hard class decisions

```{r}
safe_predict(fit, new_data = test, type = "class")
```

We can also get predictions on the link scale:


```{r}
safe_predict(fit, new_data = test, type = "link")
```

or we can get confidence intervals on the response scale

```{r}
safe_predict(fit, new_data = test, type = "conf_int")
```

## Conventions

`safepredict` is based on a conventions document that is currently in progress and private. Once conventions are finalized, they'll appear here.

