---
title: "Handling novel factor levels"
author: "Alex Hayes"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Handling novel factor levels}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## set up some example data

## warning: dummy coding variables *results in information loss*

random forest example where the wrong type of stuff happens because the factors are now treated as integers

if the model you want to work *can* deal with new factors, please let it!

```{r}
# lme4 example dealing with new factors
# brms example dealing with new factors
```

if you want to test if method x will explode, do something like this:

```{r}
# small data test for breakage
```

...here we see that things have broken. what are our options?

## options

- convert to mode
- likelihood encoding -- especially for *huge* numbers of factors
- one hot encodings -- linearly dependent columns, so lots of GLM / unpenalized likelihood methods are to break, also you'll get identifiability issues
- missing data if the prediction method can still make predictions

- dummy vignette also says you can use integer encoding, or `step_other`. don't recommend integer encoding. example of how step other might work.



## one hot encodings with recipes

### what happens to novel factor levels

For additional details, please read the full [vignette on dummy variables][dummy] in the `recipes` package.

[dummy]: https://tidymodels.github.io/recipes/articles/Dummies.html
